### 2022-01-18 plos_latex_template.tex
Just in case, pushing latest changes from Janani

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Ten simple rules towards hosting an inclusive conference} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Rocío Joo\textsuperscript{1*}, %,2\Yinyang},
Andrea Sánchez-Tapia\textsuperscript{2},
Sara Mortara\textsuperscript{3},
Yanina Bellini Saibene\textsuperscript{4},
Heather Turner\textsuperscript{5}, %\ddag},
Dorothea Hug Peter\textsuperscript{6}, %\ddag},
Matt Bannert\textsuperscript{7},
Batool Almazrouq\textsuperscript{8},
Liz Hare\textsuperscript{9},
Laura Ación\textsuperscript{10},
Juan Pablo Narváez\textsuperscript{11},
Marcela Alfaro Córdoba\textsuperscript{12},
Federico Marini\textsuperscript{13},
Rita Giordano\textsuperscript{14},
Natalia Morandeira\textsuperscript{15},
Adithi Upadhya\textsuperscript{16},
Joselyn Chávez\textsuperscript{17\Yinyang},
Anicet Ebou\textsuperscript{18\Yinyang},
Silvia Canelón\textsuperscript{19\Yinyang},
Janani Ravi\textsuperscript{20}
%,3\textcurrency}
\\
\bigskip
\textbf{1} Global Fishing Watch, Washington, DC 20036, USA
\\
\textbf{2} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{3} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{4} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{5} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{6} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{7} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{8} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{9} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{10} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{11} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{12} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{13} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{14} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{15} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{16} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{17} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{18} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{19} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{20} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\bigskip

@@

The authors of this article recently organized useR! 2021, the annual global R conference, attracting a broad range of participants from academia, industry, government, and the non-profit sector. To embrace the constraints posed by the pandemic, this year we hosted a 100\% virtual conference. We worked on building a high-quality virtual and explicitly global conference by creating a kind, inclusive, accessible, and welcoming environment for everyone. 
This article outlines our most important lessons learned abstracted as '10 simple rules to host an inclusive conference'. These rules apply equally to academic, industry, or mixed conferences; while inspired by a global experience, these rules seamlessly translate to regional events. These rules were implemented carefully and intentionally during a virtual conference, but are designed to serve as useful guidelines for in-person and hybrid events even after the COVID pandemic.

@@

% Rocío: main idea: Inclusion for everyone
While it is essential to improve representation towards some of the most visible dimensions of human diversity, such as race, gender, and country of origin, building a genuinely inclusive environment means taking care of all aspects of diversity. Inclusivity should lie at the heart of setting the tone and meeting space for any meeting as we move forward in the 21st century!
% Edit and make more specific (below)

\emoji{bulb} \textbf{Practical tips:} Here are a few ways in which conference organizers can actively enforce inclusivity, such as i) Promote respect of pronouns, ii) schedule events exclusive for diverse groups (\textit{e.g.}, LGBTQIA+-friendly spaces), iii) if the conference has an in-person component, pre-assign spaces for religious practices, lactation, and child care, iv) include menu options that account for diverse dietary and religious requirements. These are just a few easy-to-implement examples with a significant impact on including key sub-groups that otherwise find it challenging (see \cite{noauthor_discover2021} for some great advice on 'low-hanging fruits').
Importantly, by being proactive in creating such a welcoming space, the organizers can be respectful of attendees' privacy and avoid requiring anyone to disclose personal information, such as revealing a disability, sexual orientation, gender identity, or mental health issues, to a name a few.
% Rocío: Could have a paragraph about active bystandership—Andrea%ö
%promote a supportive and respectful environment where attendants act as active bystanders. y definir '¬¬ 
% Rocío: main idea: Code of conduct for a safe place

Adopting a Code of Conduct (CoC) and gathering a diverse team to enforce these ideas are vital in creating a safe environment during a conference \cite{favaroYourScienceConference2016}.
The CoC is a living document meant to keep the community safe and should state clearly the unacceptable behaviors, the consequences for engaging in such behavior, and the way to report violations \cite{auroraHowRespondCode2019}. 
The CoC team must receive training on how to receive reports, respond to incidents, communicate their responses, and organize protocols on how to respond correspondingly. This should be done during the early phases of the conference planning because CoC violations can happen before the conference too (\textit{e.g.}, social media, organization meetings).  
Since people from marginalized groups are more likely to be victims of Code of Conduct violations and find their claims dismissed due to inadvertent power dynamics, assembling a diverse CoC team will be essential to appropriately deal with discrimination and harassment issues and the power dynamics underneath them with respect and sensitivity. A well-drafted CoC may also be more understanding and accommodating of different disciplinary cultures and geographical and cultural considerations. 
We strongly recommend reading `How to Respond to Code of Conduct Reports' \cite{auroraHowRespondCode2019} as an excellent starting point for building a strong CoC team.

@@

These rules concern diverse elements in the design of the conference and operational tasks. They are all important pieces for inclusion.
In rule 4, we introduce ways to counteract bias in the conference program, in keynotes, program committee, abstract selection, and other spaces. 
Rule 5 describes the advantages of online conferences reducing barriers for participation, and provides advice for organizing online and hybrid conferences.
Rule 6 focuses on practices to make all conference platforms and spaces accessible to people with disabilities. 
In rule 7, we advocate for more language-inclusive conferences, and provide suggestions to encourage and ensure full participation of non-native English speakers. 
Rule 8 provides tips for an inclusive media strategy. 
In rule 9, we address budgeting for inclusive practices, and helping participants with affordable registration costs, scholarships, and other forms of financial support.

@@

The final list of accepted talks can also be biased in both submission and acceptance processes. 
People who have never participated in conferences, and especially those regarded as big and prestigious, may not feel confident to submit their work.
To counteract this self-selection bias, promote the submission call beyond the communication channels and reach out to local groups and communities of practice to encourage submissions by their members.
Some of these communities may have pre-submission mechanisms to help their members prepare and improve their submissions (e.g. the Latin-R community has a Slack channel with the purpose of helping community members improve their abstracts for conferences, and the R-Ladies community did the same for useR! 2021).
On the other hand, the review process may be biased because selection committees—even diverse ones—can be subjected to unconscious or implicit bias.
Unconscious bias happens when we interpret information, such as names of the authors, we assume their origin or ethnicity, their institution of origin, and we attach a positive or negative value to their work according to negative or positive stereotypes. Nobody is free from implicit bias [cite Ross2019].
Selection committees need to be reminded of such biases (the implicit association test can help identifying them [cite Greenwald - Measuring individual differences in implicit cognition: The implicit association test]), and
decide on the most appropriate strategy to counteract them, either with double blind, double open, or single blind reviews [cite discover cookbook].
Double blind reviews can help counteract implicit bias only if it is possible to completely anonymize the identifying information from submission. 
Double open, though more transparent, may still be subjected to implicit bias, and reviewers from minoritized groups might feel additional pressure because the reviews are signed. If this likely case, and in fields were it is not possible to guarantee anonymity of the authors during the selection process (e.g.  non-anonymized github repositories need to be examined), single blind reviews are the only option left.  
In any case, create clear evaluation criteria for the submitted proposals that do not put people from minoritized groups at disadvantage (e.g. based on their number of publications or talks), and share them with the authors and reviewers.
Advertising your practices to mitigate bias in proposal selection may also encourage people from diverse groups to submit their work, thus counteracting self selection bias.

>>>>

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Ten simple rules to host an inclusive conference} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Rocío Joo\textsuperscript{1*}, %,2\Yinyang},
Andrea Sánchez-Tapia\textsuperscript{2},
Sara Mortara\textsuperscript{3},
Yanina Bellini Saibene\textsuperscript{4},
Heather Turner\textsuperscript{5}, %\ddag},
Dorothea Hug Peter\textsuperscript{6}, %\ddag},
Matt Bannert\textsuperscript{7},
Batool Almazrouq\textsuperscript{8},
Liz Hare\textsuperscript{9},
Laura Ación\textsuperscript{10},
Juan Pablo Narváez\textsuperscript{11},
Marcela Alfaro Córdoba\textsuperscript{12},
Federico Marini\textsuperscript{13},
Rita Giordano\textsuperscript{14},
Natalia Morandeira\textsuperscript{15},
Adithi Upadhya\textsuperscript{16},
Joselyn Chávez\textsuperscript{17\Yinyang},
Anicet Ebou\textsuperscript{18\Yinyang},
Silvia Canelón\textsuperscript{19\Yinyang},
Janani Ravi\textsuperscript{20}
%,3\textcurrency}
\\
\bigskip
\textbf{1} Global Fishing Watch, Washington, DC 20036, USA
\\
\textbf{2} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{3} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{4} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{5} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{6} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{7} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{8} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{9} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{10} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{11} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{12} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{13} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{14} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{15} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{16} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{17} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{18} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{19} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{20} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\bigskip

@@

% % Batool: I didn't change anything in the abstract - you can check the history - I copied the abstract and pasted, then the colour change for the abstract but it contains all text and suggestions/edits by Janani - my apologies (didn't commit the changes).  
The authors of this article recently organized useR! 2021, the annual global R conference, attracting a broad range of participants from academia, industry, government, and the non-profit sector. To embrace the constraints posed by the pandemic, this year we hosted a 100\% virtual conference. We worked on building a high-quality virtual and explicitly global conference by creating a kind, inclusive, accessible, and welcoming environment for everyone. 
This article outlines our most important lessons learned abstracted as '10 simple rules to host an inclusive conference'. These rules apply equally to academic, industry, or mixed conferences; while inspired by a global experience, these rules seamlessly translate to regional events. These rules were implemented carefully and intentionally during a virtual conference, but are designed to serve as useful guidelines for in-person and hybrid events even after the COVID pandemic.

@@

% Rocío: main idea: Inclusion for everyone
While it is common to think about representation in terms of some of the most visible dimensions of human diversity, such as race, gender, and country of origin, building a genuinely inclusive environment means taking care of all aspects of diversity. Inclusivity should lie at the heart of setting the tone and meeting space for any meeting as we move forward in the 21st century!
% Edit and make more specific (below)

\emoji{bulb} \textbf{Practical tips:} Here are a few ways in which conference organizers can actively enforce inclusivity, such as i) Promote respect of pronouns, ii) schedule events exclusive for diverse groups (\textit{e.g.}, LGBTQIA+-friendly spaces), iii) if the conference has an in-person component, pre-assign spaces for religious practices, lactation, and child care, iv) include menu options that account for diverse dietary and religious requirements. These are just a few easy-to-implement examples with a significant impact on including key sub-groups that otherwise find it challenging (see \cite{noauthor_discover2021} for some great advice on 'low-hanging fruits').
Importantly, by being proactive in creating such a welcoming space, the organizers can be respectful of attendees' privacy and avoid requiring anyone to disclose personal information, such as revealing a disability, sexual orientation, gender identity, or mental health issues, to a name a few.

The organizers should promote a supportive and respectful environment, where all attendants can be themselves without fear of rejection or isolation. This includes promoting the idea that anyone can act in defense of other attendants in case anything unpleasant, incorrect, of offensive happens. Instead of looking away, act as "active bystanders" and intervene to keep the community safe.
[ast: should we add pre-conference allyship workshops? i saw a conference that had one pre-conference events for the whole community, just brilliant]
% Rocío: main idea: Code of conduct for a safe place
However, the most important step that organizers should take to ensure the safety of the community is
Adopting a Code of Conduct (CoC) and gathering a diverse team to enforce it \cite{favaroYourScienceConference2016}.
The CoC is a living document meant to keep the community safe and should state clearly the unacceptable behaviors, the consequences for engaging in such behavior, and the way to report violations \cite{auroraHowRespondCode2019}. 
The CoC team must receive training on how to receive reports, respond to incidents, communicate their responses, and organize protocols on how to respond correspondingly. This should be done during the early phases of the conference planning because CoC violations can happen before the conference too (\textit{e.g.}, social media, organization meetings).  
Since people from marginalized groups are more likely to be victims of Code of Conduct violations and find their claims dismissed due to inadvertent power dynamics, assembling a diverse CoC team will be essential to appropriately deal with discrimination and harassment issues and the power dynamics underneath them with respect and sensitivity. A well-drafted CoC may also be more understanding and accommodating of different disciplinary cultures and geographical and cultural considerations. 
We strongly recommend reading `How to Respond to Code of Conduct Reports' \cite{auroraHowRespondCode2019} as an excellent starting point for building a strong CoC team.

@@

The next set of rules concern diverse elements in the design of the conference and operational tasks. They focus on weaving inclusion into the conference design process.
In \textbf{Rule 4}, we introduce ways to counteract bias in the conference program, keynotes, program committee, abstract selection, and other spaces. 
\textbf{Rule 5} describes the advantages of online conferences by reducing barriers for participation and provides advice for organizing online and hybrid conferences.
\textbf{Rule 6} focuses on best practices to make all conference platforms and spaces accessible to people with disabilities. 
In \textbf{Rule 7}, we advocate for more language-inclusive conferences and provide suggestions to encourage and ensure full participation of non-native English speakers. 
\textbf{Rule 8} provides tips for developing an inclusive media strategy. 
In \textbf{Rule 9}, we address budgeting for inclusive practices, and helping participants with affordable registration costs, scholarships, and other forms of financial support.

@@

The final list of accepted talks can also be biased in both submission and acceptance processes. 
People who have never participated in conferences, and especially those regarded as big and prestigious, may not feel confident to submit their work.
To counteract this self-selection bias, promote the submission call beyond the communication channels and reach out to local groups and communities of practice to encourage submissions by their members.
Some of these communities may have pre-submission mechanisms to help their members prepare and improve their submissions (e.g. the Latin-R community has a Slack channel with the purpose of helping community members improve their abstracts for conferences, and the R-Ladies community did the same for useR! 2021).
On the other hand, the review process may be biased because selection committees—even diverse ones—can be subjected to unconscious or implicit bias.
Unconscious bias happens when we interpret information, such as names of the authors, we assume their origin or ethnicity, their institution of origin, and we attach a positive or negative value to their work according to negative or positive stereotypes. Nobody is free from implicit bias [cite Ross2019].
Selection committees need to be reminded of such biases (the implicit association test can help identifying them [cite Greenwald - Measuring individual differences in implicit cognition: The implicit association test]), and
decide on the most appropriate strategy to counteract them, either with double blind, double open, or single blind reviews [cite discover cookbook].
Double blind reviews can help counteract implicit bias only if it is possible to completely anonymize the identifying information from submission. 
Double open, though more transparent, may still be subjected to implicit bias, and reviewers from minoritized groups might feel additional pressure because the reviews are signed. If this likely case, and in fields were it is not possible to guarantee anonymity of the authors during the selection process (e.g.  non-anonymized GitHub repositories need to be examined), single blind reviews are the only option left.  
In any case, create clear evaluation criteria for the submitted proposals that do not put people from minoritized groups at disadvantage (e.g. based on their number of publications or talks), and share them with the authors and reviewers.
Advertising your practices to mitigate bias in proposal selection may also encourage people from diverse groups to submit their work, thus counteracting self selection bias.
